---
title: "Up and Running with R"
author: "YOUR NAME"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading Notes

Replace this text with your reading notes. You must include at least three new concepts that you learned.

You can try making your text **bold** or *italicized* to help organize your notes.

You can also include bullet points:

- Like this
- And this

If you'd like to learn more about formatting your text using Markdown, you can use [this reference](https://www.markdownguide.org/basic-syntax/).

## R Practice

### Question 1

Create an object named `object1` and assign it the value of `10`. Create another object named `object2` and assign it the value of `"BYU-Idaho"`. Then try to add the two values together.

**What error message do you get?** Type your answer here.

 - Error in object1 + object2: non-numeric argument to binary operator

**What is this error message trying to tell you?** Type your answer here.

 - R treated object1 + object2 as a boolean statement to compare if both objects are TRUE and also complained that object2 was not a number so simple addition won't work either.

```{r, error = TRUE}
object1 <- 10
object2 <- 'BYU-idaho'
object1 + object2
# x %>% f is pipe which means f(x)
# f(x,y,z) is same as x %>% f(y,z)
# View, Glimpse, Head, Tail
```
### Question 2

Write a function that takes two numbers as input and returns their mean. (Do not use the built in `mean()` function.) Use your function to assign the mean of 28 and 14 to an object called `result`. Display `result`.

```{r}
func1 <- function(num1, num2)
  return((num1 + num2) / 2)
num1 <- 28
num2 <- 14
func1(num1, num2)
```

### Question 3

Create a vector that contains the temperatures 85, 83, 83, 89, 92, 87, and 84. Then calculate:

- The mean of the temperatures
- The standard deviation of the temperatures
- The deviation of each temperature from the mean temperature 

```{r}

vector <- c(85, 83, 83, 89, 92, 87, 84)
mean <- mean(vector)
sd <- sd(vector)
abs((vector-mean) / sd)
```

### Question 4

Create a vector `v` with all integers 0-30, and a vector `w` with every third integer in the same range. (The documentation for `?seq` might be helpful.)

Calculate the difference in lengths of the vectors `v` and `w`. 

```{r}
v <- seq(from=0, to=30)
w <- seq(from=0,to=30, by=3)
length(v) - length(w)
```

### Question 5

Create a new vector `v_square` with the square of elements at indices 3, 6, 7, 10, 15, 22, 23, 24, and 30 from the variable `v`. Display `v_square`.

```{r}
v_square <- v[c(3,6,7,10,15,22,23,24,30)]^2
v_square
```

### Question 6

Create a boolean vector `v_bool` indicating which vector `v` elements are bigger than 20. Display `v_bool`. Then calculate how many values are over 20. (Hint: In R, TRUE = 1, and FALSE = 0, so you can use simple arithmetic to find this out.)

```{r}
v_bool <- v > 20
length(v[v_bool])
```

### Question 7

Display the output of `v[TRUE]`. 

**Explain why you think R outputs this.** Type your answer here. 
 
 In the previous problem, the answer paired each boolean value with the vector v value in the same position. Since we are giving a vector of values of length 1, that is smaller than the actual length, it repeats the values given. Since it's only TRUE, it repeats TRUE for every value.
```{r}
v[TRUE]
```

### Question 8

Use the variable `v_bool` as an index to extract the elements from `v` that are bigger than 20. Calculate the min and max values of this new vector.

```{r}
new_v = v[v[v_bool]]
min(new_v)
max(new_v)
```

### Question 9

What are the column names of the built-in dataframe `beaver1`? How many observations (rows) and variables (columns) are there?

```{r}
colnames(beaver1)
num_col = length(beaver1)
num_len = length(t(beaver1))
num_col
num_len
```

### Question 10

Use square bracket notation to index the `beaver1` data set. Display the values in rows 21-23, columns 3-4.

```{r}
beaver1[21:23,3:4]
```

### Question 11

What is the min, mean, and max body temperature in the `beaver1` data set?

```{r}
min(beaver1$temp)
max(beaver1$temp)
```

### Question 12

In the `beaver1` data set the variable `temp` is recorded in degrees Celsius. Run the code below, which is trying to convert `temp` to Fahrenheit.

**What error message do you get?** Type your answer here.

 - Error in c_to_f(beaverl$temp) : object 'beaverl' not found

**What is this error message trying to tell you?** Type your answer here.

That there is an l in beaverl instead of a 1 for beaver1.

```{r, error = TRUE}
# function to convert Celsius to Fahrenheit
c_to_f <- function(c){
  f = (c * 9/5) + 32
  return(f)
}

# make a new vector of Fahrenheit temperatures!
temp_in_f <- c_to_f(beaverl$temp)
```