---
title: "task_7"
output: md_document
---

Notes:
You can use filter to filter the rows. Like Where in SQL.
Comparitors are <, >, ==, and %in%
And is &
Or is |
not is !

Use is.na(x) to test if a value is not present
You can rename column names with the rename function
select is a query
You can create new columns through mutate and multiple at a time separated by a comma
logarithms are used through log(), log2(), log10()
lead and lag can be used to sselect a range specified ie 3:10 with an index from either start or finish
min_rank() can be used for ranking values in order
summarise(mean = mean(dep_delay, na.rm = TRUE))



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remotes::install_github("apreshill/bakeoff")
library(bakeoff)
library(tidyverse)
```

## R Markdown

```{r}
only_first_last <- ratings %>% 
  group_by(series) %>% 
  slice(1,n()) %>% 
  mutate(which_episode = ifelse(episode==1, "First", "Last")) %>% 
  ungroup() %>% 
  mutate(series_f = as.factor(series))

only_first_last %>% glimpse()


only_first_last %>% 
  ggplot(mapping=aes(x=which_episode,
                     y=viewers_7day,
                     group=series_f,
                     color=series_f)) +
  geom_line() +
  geom_point(size=5)
```


```{r}
d<-ratings %>% 
  group_by(series) %>% 
  slice_tail(n=1) %>% 
  filter(episode %in% 10)
epi_10_season <- ratings %>% 
  group_by(series) %>% 
  filter(series %in% d$series)

mean_epi<-aggregate(. ~ epi_10_season$episode, epi_10_season[4], mean) %>% 
  rename(
    episode=`epi_10_season$episode`
  )
epi_10_season <- epi_10_season %>% inner_join(mean_epi, by="episode")
epi_10_season %>% 
  ggplot() +
  geom_line(mapping=aes(x=episode, y=viewers_7day.x, group=series, color=series)) +
  geom_point(mapping=aes(x=episode, y=viewers_7day.x, group=series, color=series), size=3)
```




```{r}
par(mfrow=c(2,1))
epi_10_season %>% 
  ggplot() +
  geom_line(mapping=aes(x=episode, y=viewers_7day.x, group=series, color=series)) +
  geom_point(mapping=aes(x=episode, y=viewers_7day.x, group=series, color=series), size=3)
epi_10_season %>% 
  ggplot() +
  geom_col(mapping=aes(x=episode, y=epi_10_season$viewers_7day.y))
```


